Parallel air palletization and tour planning for simultaneous pickup and delivery

### 2022-11-15 latex/comp_or.tex
ACO_p experiments completed

In the scenario of aerial pickup and delivery of goods in a distribution network, transport aviation faces risks of cargo unbalancing due to the urgency required for loading for rapid take-off and mission accomplishment, especially in times of crisis, public calamity, client contract short lead times, or any external pressure for immediate take-off. Also, there is no technological assistance to help the load and trip planners with the huge number of demands for transport in each hub.
This enables other risks such as improper delivery, excessive fuel burn, and longer than necessary turn-around time. We contributed by modeling and solving a new problem of planning the loading and routing of an aircraft according to a utility score, weight and balance principles, and fuel consumption in a tour of simultaneous pickup and delivery at intermediate hubs.
This new hard problem, named {\it Air Cargo Load Planning with Routing, Pickup, and Delivery Problem (ACLP+RPDP)}, is mathematically modeled using standardized pallets in fixed positions, obeying the center of gravity constraints, delivering each item to its destination, and minimizing fuel consumption costs.
We also contributed by carrying out multiple experiments with the well-known Ant Colony Optimization on synthetic data based on real data from the {\it Brazilian Air Force}\/ transportation history.
These challenging benchmark instances are made publicly available. We also created a multiprocessing computing heuristic that quickly finds good solutions for a wide range of problem sizes, an essential contribution as it was the unique method that managed to solve all testing scenarios.

@@

However, there are still other important challenges in air cargo transport that go beyond the definition of the ACLPP, especially with regard to the flight itinerary and the loading and unloading at each destination (or node) of this travel plan. In this context, at least two more important sub-problems can be considered: pickup and delivery operations at each node, called {\it Pickup and Delivery Problem} (PDP), and the search for the lowest cost route, which is the well-known {\it Traveling Salesman Problem} (TSP).

@@

\begin{table}[H]
	\centering
	\caption{Air cargo transport: literature, problems and features}  \label{tab:sa}
	\scriptsize
	\renewcommand{\arraystretch}{1.1} % alturas das linhas
	\begin{tabular}{r|cc|cc|cc|ccc}
		\toprule
		                            & {\bf APP}  & {\bf WBP}  &  {\bf PDP}   &{\bf TSP}   & {\bf 2D}  & {\bf 3D}  & {\bf Heu}  & {\bf Int}  & {\bf Lin} \\
		\midrule
		\cite{LarsenMikkelsen1979}  & $.$        & $\bigstar$ & $.$          & $.$        & $.$       & $.$       & $\bigstar$ & $.$        &  $.$ \\
		\cite{Brosh1981}  & $.$ & $\bigstar$  & $.$   & $.$ & $.$ & $.$   & $.$  & $.$  &  $\bigstar$ \\
		\cite{Kevin1992}  & $.$ & $\bigstar$  & $.$ & $.$ & $.$ &$.$   & $.$  & $\bigstar$  &  $.$ \\
		\cite{Heidelberg1998}  & $.$ & $\bigstar$  & $.$ & $.$ & $\bigstar$ &$.$   & $\bigstar$  & $.$  &  $.$ \\
		\cite{MongeauBes2003}    & $\bigstar$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\cite{fok2004optimizing} & $.$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\	
		\cite{Chan2006}  & $\bigstar$ & $.$    & $.$ & $.$ & $.$ & $\bigstar$  & $\bigstar$  & $.$  &  $.$ \\
		\cite{KaluznyBohdanL2009Oalb}  & $.$ & $\bigstar$  & $.$  & $.$ & $\bigstar$ &$.$  & $.$  & $\bigstar$  &  $.$ \\
		\cite{Verstichel2011}   & $.$ & $\bigstar$    & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\cite{MesquitaCunha2011}   & $.$ & $.$    & $\bigstar$ & $.$ & $.$ & $.$   & $\bigstar$  & $.$  &  $.$ \\		
		\cite{Limbourg2012} & $.$ & $\bigstar$  & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\cite{RoesenerHall2014}  & $\bigstar$ & $\bigstar$  & $.$  & $.$ & $.$ & $\bigstar$   & $.$  & $\bigstar$  &  $.$ \\
		\cite{Vancroonemburg2014}  & $\bigstar$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\cite{LurkinSchyns2015} & $.$ & $\bigstar$  & $\bigstar$ & $.$  & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\cite{RoesenerBarnes2016}  & $.$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $\bigstar$  & $.$  &  $.$ \\
		\cite{PaquaySchynsLimbourg2016,PaquayLimbourgSchynsOliveira2018}  & $\bigstar$ & $\bigstar$ & $.$ & $.$ & $.$ & $\bigstar$ & $\bigstar$  & $\bigstar$ & $.$ \\
		\cite{YangLiuGao2018} & $.$ & $\bigstar$  & $.$  & $.$ & $\bigstar$  & $.$ & $\bigstar$ & $.$  & $.$ \\
		\cite{wong2020} & $\bigstar$  & $\bigstar$  & $.$  & $.$   & $.$  & $.$ & $.$ & $\bigstar$  & $.$  \\
		\cite{eugene2021} & $\bigstar$ & $\bigstar$ & $.$  & $.$   & $.$ & $.$ & $.$ & $\bigstar$  & $.$  \\
		\cite{zhao2021} & $.$ & $\bigstar$ & $.$  & $.$  & $.$ & $.$ & $.$  & $\bigstar$ &  $.$ \\
		{\bf This work}   & $\bigstar$ & $\bigstar$  & $\bigstar$& $\bigstar$ & $.$ & $.$ & $\bigstar$ & $\bigstar$   &  $.$  \\
		\bottomrule 
	\end{tabular}
	\normalsize 
\end{table}

@@

@@

As already mentioned, \cite{LurkinSchyns2015} was the first work that simultaneously modeled WBP and PDP in air cargo transport. The authors demonstrated that this problem is NP-hard and performed some experiments with real data, noting that their model offers better results than those obtained manually.

@@

As can be seen, none of these works address air cargo palletization and load balancing with route optimization in a multi-leg flight plan. This is the objective of our work: to model and elaborate heuristics for a real problem that simultaneously involves 4 intractable sub-problems: APP, WBP, PDP and TSP.

@@

	\item We considered a unique pallet type: the {\it 463L Master Pallet}, a common size platform for bundling and moving air cargo. It is the primary air cargo pallet for more than 70 Air Forces and many air transport companies. This pallet has a capacity of $4500 kg$ and $13.7 m^3$, is equipped for locking into cargo aircraft rail systems, and includes tie-down rings to secure nets and cargo loads, which in total weighs $140 kg$. For more information, see {\tt www.463LPallet.com}.

@@

Real cases are more complex as they have hundreds of different items in each node and involve three intractable sub-problems: APP, WBP and PDP. Through the mathematical modeling presented in the previous section, we verify that {\it Mixed-Integer Programming}\/ (MIP) is not able to solve these cases in feasible time. Thus, it is necessary to adopt some strategy to find a viable solution, not necessarily optimal, that seeks to maximize the objective function $f$.

@@

	\begin{algorithmic}[1]
		\State Let $G(V_k, E^N_k \cup E^Q_k)$
		\State Let $X_{ij}^k = 1$ if $(p_i, t_j^k) \in E^N_k$, $1 \leq i \leq m$, $1 \leq j \leq n_k$
		\State Let $Y_{iq}^k = 1$ if $(p_i, a_q^k) \in E^Q_k$, $1 \leq i \leq m$, $1 \leq q \leq m_k$
		\State $s \gets 0$
		\State $\epsilon \gets 0$
		\State $weight \gets 0$
		\For{$i \gets 1$ to $m$}	\label{eval:loop1}
			\State $weight \gets weight + p_i.w$		
			\For{$j \gets 1$ to $n_k$}
				\If{$X_{ij}^k = 1$}
					\State $s \gets s + t_{ij}^k.s$ \label{eval:score1}
					\State $\epsilon \gets \epsilon + t_{ij}^k.w * p_i.d$ \label{eval:eps1}
				\EndIf
			\EndFor				
			\For{$q \gets 1$ to $m_k$}
				\If{$Y_{iq}^k = 1$}
					\State $s \gets s + a_{iq}^k.s$   \label{eval:score2}
					\State $\epsilon \gets \epsilon + a_{iq}^k.w * p_i.d$  \label{eval:eps2}
				\EndIf
			\EndFor	   \label{eval:loop2}
		\EndFor
		\State $maxWeight \gets \min(Payload, weight)$ \label{eval:torque1}
		\State $maxTorque \gets maxWeight \times limit^{CG}_{long}$
		\State $\epsilon \gets \epsilon/maxTorque$ \label{eval:torque2}
		\State ${\bf return}\ s, \epsilon$	
	\end{algorithmic}
\end{algorithm}

@@

An important issue on meta-heuristics is that they are known for the hard work in tuning parameters. For this, when we use the word {\it empirically}\/ throughout the text, we want to say that we accomplished many tests with the hardest instances to fit these values. This does not guarantee best results for all instances, but, on average, overall results are acceptable. Another important observation is that our heuristics were designed based on traditional ideas from the literature \cite{Fidanova2006} and on the experiences learned during tests.

@@

All the experiments are performed on a 64-bit, 128GiB, 3.6GHz, 48-core processor, 2 threads per core, Intel Xeon, with {\it Linux CentOS 8} as the operational system and {\it Python 3.10.4}\/ as the programming language.

@@

The {\it Multiprocessing} package (mp) offers both local and remote concurrency, effectively side-stepping the global interpreter lock by using sub-processes instead of threads. Because of this, the {\it Multiprocessing} module lets the programmer take full advantage of the fact that a machine has more than one processor, 48 in our computer. This decision for {\it Multiprocessing} relies on the need to exchange information among processes (the pheromone trails or some information regarding load balancing, in our case), which would be harder to handle with thread-based parallelism.

@@

\begin{figure}[H]
	\begin{center}
		\begin{tikzpicture}[scale=0.8, samples=100]
			\begin{axis}[
				myplotstyle,
				xlabel={\bf Scenario 1 computation times (s)},
				ylabel={\bf Score},
				]
				\addplot+[
				only marks,	
				nodes near coords = {\numProcs{}},
				visualization depends on = {value \thisrow{np} \as \numProcs},						
				]
				table[x=time, y=score, col sep=comma] {./csv/Shims_p1data20.csv};

				\legend{\it parallel processes}
			\end{axis}
		\end{tikzpicture}%
		~%
		%
		\begin{tikzpicture}[scale=0.8, samples=100]
			\begin{axis}[
				myplotstyle,
				xlabel={\bf Scenario 2 computation times (s)},
				]
				\addplot+[
				only marks,	
				nodes near coords = {\numProcs{}},
				visualization depends on = {value \thisrow{np} \as \numProcs},						
				]
				table[x=time, y=score, col sep=comma] {./csv/Shims_p2data20.csv};

				\legend{\it parallel processes}
			\end{axis}
		\end{tikzpicture}

@@

Observing Figure \ref{fig:mpShimsdata20}, it may be noticed that the best number of processes for {\it Scenario 1} (closest to the upper left corner) is \textbf{1}. For {\it Scenario 2}, the best number is \textbf{4}.

\begin{figure}[H]
	\begin{center}
		\begin{tikzpicture}[scale=0.8, samples=100]
			\begin{axis}[
				myplotstyle,
				xlabel={\bf Scenario 1 computation times (s)},
				ylabel={\bf Score},
				]
				\addplot+[
				only marks,	
				nodes near coords = {\numProcs{}},
				visualization depends on = {value \thisrow{np} \as \numProcs},						
				]
				table[x=time, y=score, col sep=comma] {./csv/Shims_p1data50.csv};

				\legend{\it parallel processes}
			\end{axis}
		\end{tikzpicture}%
		~%
		%
		\begin{tikzpicture}[scale=0.8, samples=100]
			\begin{axis}[
				myplotstyle,
				xlabel={\bf Scenario 2 computation times (s)},
				]
				\addplot+[
				only marks,	
				nodes near coords = {\numProcs{}},
				visualization depends on = {value \thisrow{np} \as \numProcs},						
				]
				table[x=time, y=score, col sep=comma] {./csv/Shims_p2data50.csv};

				\legend{\it parallel processes}
			\end{axis}
		\end{tikzpicture}

	\end{center}
	\caption{{\it mpShims} performance with $1.5 \times volume$.}
	\label{fig:mpShimsdata50}
\end{figure}

Observing Figure \ref{fig:mpShimsdata50}, it may be noticed that the best number of processes for {\it Scenario 1} is \textbf{1}. For {\it Scenario 2}, the best number is \textbf{8}.

@@

	\begin{center}
		\begin{tikzpicture}[scale=0.8, samples=100]
			\begin{axis}[
				myplotstyle,
				xlabel={\bf Scenario 1 computation times (s)},
				ylabel={\bf Score},
				]
				\addplot+[
				only marks,	
				nodes near coords = {\numProcs{}},
				visualization depends on = {value \thisrow{np} \as \numProcs},						
				]
				table[x=time, y=score, col sep=comma] {./csv/ACO_p1data20.csv};

				\legend{\it parallel processes}
			\end{axis}
		\end{tikzpicture}%
		~%
		%
		\begin{tikzpicture}[scale=0.8, samples=100]
			\begin{axis}[
				myplotstyle,
				xlabel={\bf Scenario 2 computation times (s)},
				]
				\addplot+[
				only marks,	
				nodes near coords = {\numProcs{}},
				visualization depends on = {value \thisrow{np} \as \numProcs},						
				]
				table[x=time, y=score, col sep=comma] {./csv/ACO_p2data20.csv};

				\legend{\it parallel processes}
			\end{axis}
		\end{tikzpicture}

	\end{center}
	\caption{{\it mpACO} performance with $1.2 \times volume$.}
	\label{fig:mpACOdata20}
\end{figure}

Observing Figure \ref{fig:mpACOdata20}, it may be noticed that the best number of processes for {\it Scenario 1} is \textbf{40}. For {\it Scenario 2}, the best number is \textbf{1}.

	\begin{center}
		\begin{tikzpicture}[scale=0.8, samples=100]
			\begin{axis}[
				myplotstyle,
				xlabel={\bf Scenario 1 computation times (s)},
				ylabel={\bf Score},
				]
				\addplot+[
				only marks,	
				nodes near coords = {\numProcs{}},
				visualization depends on = {value \thisrow{np} \as \numProcs},						
				]
				table[x=time, y=score, col sep=comma] {./csv/ACO_p1data50.csv};

				\legend{\it parallel processes}
			\end{axis}
		\end{tikzpicture}%
		~%
		%
		\begin{tikzpicture}[scale=0.8, samples=100]
			\begin{axis}[
				myplotstyle,
				xlabel={\bf Scenario 2 computation times (s)},
				]
				\addplot+[
				only marks,	
				nodes near coords = {\numProcs{}},
				visualization depends on = {value \thisrow{np} \as \numProcs},						
				]
				table[x=time, y=score, col sep=comma] {./csv/ACO_p2data50.csv};

				\legend{\it parallel processes}
			\end{axis}
		\end{tikzpicture}

	\end{center}
	\caption{{\it mpACO} performance with $1.5 \times volume$.}
	\label{fig:mpACOdata50}
\end{figure}

Observing Figure \ref{fig:mpACOdata50}, it may be noticed that the best number of processes for {\it Scenario 1} is \textbf{1}. For {\it Scenario 2}, the best number is \textbf{16}.

@@

		\multirow{2}{*}{\it {mpShims}} &\multirow{2}{*}{\bf {24}}& $f$              & 5.88        & 3.16    & 3.16         &          &          &         &   \\ 
		                               &                         &{\bf run-time (s)}& 1           & 10      & 41           &            &            &          &   \\ %299s
		\midrule[.1pt]

@@

        \multirow{2}{*}{\it {mpACO}}   &\multirow{2}{*}{\bf {32}}& $f$              & 5.82        & 3.12    &          &         &       &       &   \\ 
                                       &                         &{\bf run-time (s)}& 22          & 473     &           &         &       &       &   \\

>>>>

In the scenario of aerial pickup and delivery of goods in a distribution network, transport aviation faces risks of cargo unbalancing due to the urgency required for loading for rapid take-off and mission accomplishment, especially in times of crisis, public calamity, client contract short lead times, or any external pressure for immediate take-off. Also, there is no technological assistance to help the load and trip planners with the huge number of demands for transport in each hub.
This enables other risks such as improper delivery, excessive fuel burn, and longer than necessary turn-around time. We contributed by modeling and solving a new problem of planning the loading and routing of an aircraft according to a utility score, weight and balance principles, and fuel consumption in a tour of simultaneous pickup and delivery at intermediate hubs.
This new hard problem, named {\it Air Cargo Load Planning with Routing, Pickup, and Delivery Problem (ACLP+RPDP)}, is mathematically modeled using standardized pallets in fixed positions, obeying the center of gravity constraints, delivering each item to its destination, and minimizing fuel consumption costs.
We also contributed by carrying out multiple experiments with the well-known Ant Colony Optimization on synthetic data based on real data from the {\it Brazilian Air Force}\/ transportation history.
These challenging benchmark instances are made publicly available.
We also created a process-based parallel computing heuristic that quickly finds good solutions for a wide range of problem sizes, an essential contribution as { \color{red} it was the unique method that managed to solve all testing scenarios.}

@@

However, there are still other important challenges in air cargo transport that go beyond the definition of the ACLPP, especially with regard to the flight itinerary and the loading and unloading at each destination (or node) of this travel plan. In this context, at least two more important sub-problems can be considered: pickup and delivery operations at each node, called {\it Simultaneous Pickup and Delivery Problem} (SPDP), and the search for the most profitable route, which is the well-known {\it Traveling Salesman Problem} (TSP), similar to the proposed by \cite{kaspi2019}, where the profit per unit time is defined as the profit gained after completing the tour divided by the total time required to complete the tour.

@@

\begin{table}[H]
	\centering
	\caption{Air cargo transport: literature, problems and features}  \label{tab:sa}
	\scriptsize
	\renewcommand{\arraystretch}{1.1} % alturas das linhas
	\begin{tabular}{r|cc|cc|cc|ccc}
		\toprule
		                            & {\bf APP}  & {\bf WBP}  &  {\bf SPDP}   &{\bf TSP}   & {\bf 2D}  & {\bf 3D}  & {\bf Heu}  & {\bf Int}  & {\bf Lin} \\
		\midrule
		\cite{LarsenMikkelsen1979}  & $.$        & $\bigstar$ & $.$          & $.$        & $.$       & $.$       & $\bigstar$ & $.$        &  $.$ \\
		\cite{Brosh1981}  & $.$ & $\bigstar$  & $.$   & $.$ & $.$ & $.$   & $.$  & $.$  &  $\bigstar$ \\
		\cite{Kevin1992}  & $.$ & $\bigstar$  & $.$ & $.$ & $.$ &$.$   & $.$  & $\bigstar$  &  $.$ \\
		\cite{Heidelberg1998}  & $.$ & $\bigstar$  & $.$ & $.$ & $\bigstar$ &$.$   & $\bigstar$  & $.$  &  $.$ \\
		\cite{MongeauBes2003}    & $\bigstar$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\cite{fok2004optimizing} & $.$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\	
		\cite{Chan2006}  & $\bigstar$ & $.$    & $.$ & $.$ & $.$ & $\bigstar$  & $\bigstar$  & $.$  &  $.$ \\
		\cite{KaluznyBohdanL2009Oalb}  & $.$ & $\bigstar$  & $.$  & $.$ & $\bigstar$ &$.$  & $.$  & $\bigstar$  &  $.$ \\
		\cite{Verstichel2011}   & $.$ & $\bigstar$    & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\cite{MesquitaCunha2011}   & $.$ & $.$    & $\bigstar$ & $.$ & $.$ & $.$   & $\bigstar$  & $.$  &  $.$ \\		
		\cite{Limbourg2012} & $.$ & $\bigstar$  & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\

		\cite{kaspi2019} & $.$ & $.$  & $.$ & $\bigstar$ & $.$ & $.$   & $\bigstar$  & $.$  &  $.$ \\

		\cite{RoesenerHall2014}  & $\bigstar$ & $\bigstar$  & $.$  & $.$ & $.$ & $\bigstar$   & $.$  & $\bigstar$  &  $.$ \\
		\cite{Vancroonemburg2014}  & $\bigstar$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\cite{LurkinSchyns2015} & $.$ & $\bigstar$  & $\bigstar$ & $.$  & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\cite{RoesenerBarnes2016}  & $.$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $\bigstar$  & $.$  &  $.$ \\
		\cite{PaquaySchynsLimbourg2016,PaquayLimbourgSchynsOliveira2018}  & $\bigstar$ & $\bigstar$ & $.$ & $.$ & $.$ & $\bigstar$ & $\bigstar$  & $\bigstar$ & $.$ \\
		\cite{YangLiuGao2018} & $.$ & $\bigstar$  & $.$  & $.$ & $\bigstar$  & $.$ & $\bigstar$ & $.$  & $.$ \\
		\cite{wong2020} & $\bigstar$  & $\bigstar$  & $.$  & $.$   & $.$  & $.$ & $.$ & $\bigstar$  & $.$  \\
		\cite{eugene2021} & $\bigstar$ & $\bigstar$ & $.$  & $.$   & $.$ & $.$ & $.$ & $\bigstar$  & $.$  \\
		\cite{zhao2021} & $.$ & $\bigstar$ & $.$  & $.$  & $.$ & $.$ & $.$  & $\bigstar$ &  $.$ \\
		{\bf This work}   & $\bigstar$ & $\bigstar$  & $\bigstar$& $\bigstar$ & $.$ & $.$ & $\bigstar$ & $\bigstar$   &  $.$  \\
		\bottomrule 
	\end{tabular}
	\normalsize 
\end{table}

@@

\cite{kaspi2019} define and solve a new extension of the TSP to maximize the financial contribution per invested time and present an optimal iterative solution procedure for the problem which converges after a limited number of iterations.

@@

As already mentioned, \cite{LurkinSchyns2015} was the first work that simultaneously modeled WBP and SPDP in air cargo transport. The authors demonstrated that this problem is NP-hard and performed some experiments with real data, noting that their model offers better results than those obtained manually.

@@

As can be seen, none of these works address air cargo palletization and load balancing with route optimization in a multi-leg flight plan. This is the objective of our work: to model and elaborate heuristics for a real problem that simultaneously involves 4 intractable sub-problems: APP, WBP, SPDP and TSP.

@@

	\item We considered a unique pallet type: the {\it 463L Master Pallet}, a common size platform for bundling and moving air cargo. It is the primary air cargo pallet for more than 70 Air Forces and many air transport companies. This pallet has a capacity of $4500 kg$ and $14.8 m^3$, is equipped for locking into cargo aircraft rail systems, and includes tie-down rings to secure nets and cargo loads, which in total weighs $140 kg$. For more information, see {\tt www.463LPallet.com}.

@@

Real cases are more complex as they have hundreds of different items in each node and involve three intractable sub-problems: APP, WBP and SPDP. Through the mathematical modeling presented in the previous section, we verify that {\it Mixed-Integer Programming}\/ (MIP) is not able to solve these cases in feasible time. Thus, it is necessary to adopt some strategy to find a viable solution, not necessarily optimal, that seeks to maximize the objective function $f$.

@@

	\begin{algorithmic}[1]
		\State Let $G(V_k, E^N_k \cup E^Q_k)$
		\State Let $X_{ij}^k = 1$ if $(p_i, t_j^k) \in E^N_k$, $1 \leq i \leq m$, $1 \leq j \leq n_k$
		\State Let $Y_{iq}^k = 1$ if $(p_i, a_q^k) \in E^Q_k$, $1 \leq i \leq m$, $1 \leq q \leq m_k$
		\State $s \gets 0$
		\State $\tau \gets 0$
		\State $weight \gets 0$
		\For{$i \gets 1$ to $m$}	\label{eval:loop1}
			\State $weight \gets weight + p_i.w$		
			\For{$j \gets 1$ to $n_k$}
				\If{$X_{ij}^k = 1$}
					\State $s \gets s + t_{ij}^k.s$ \label{eval:score1}
					\State $\tau \gets \tau + t_{ij}^k.w * p_i.d$ \label{eval:eps1}
				\EndIf
			\EndFor				
			\For{$q \gets 1$ to $m_k$}
				\If{$Y_{iq}^k = 1$}
					\State $s \gets s + a_{iq}^k.s$   \label{eval:score2}
					\State $\tau \gets \tau + a_{iq}^k.w * p_i.d$  \label{eval:eps2}
				\EndIf
			\EndFor	   \label{eval:loop2}
		\EndFor
		\State $maxWeight \gets \min(Payload, weight)$ \label{eval:torque1}
		\State $maxTorque \gets maxWeight \times limit^{CG}_{long}$
		\State $\epsilon \gets \tau/maxTorque$ \label{eval:torque2}
		\State ${\bf return}\ s, \epsilon$	
	\end{algorithmic}
\end{algorithm}

@@

An important issue on meta-heuristics is that they are known for the hard work in tuning parameters. For this, when we use the word {\it empirically}\/ throughout the text, we want to say that we accomplished many tests with the hardest instances to fit these values. This does not guarantee best results for all instances, but, on average, overall results are acceptable. Another important observation is that our {\it ACO} implementations were designed based on traditional ideas from the literature \cite{Fidanova2006} and on the experiences learned during tests, plus the special features to process-based parallelism.

@@

%All the experiments are performed on a 64-bit, 128GiB, 3.6GHz, 48-core processor, 2 threads per core, Intel Xeon, with {\it Linux CentOS 8} as the operational system and {\it Python 3.10.4}\/ as the programming language.

All the experiments are performed on a 64-bit, 16GiB, 3.2GHz, 8-core processor, 2 threads per core, AMD FX-8320E, with {\it Linux Ubuntu 22.04} as the operational system and {\it Python 3.10.4}\/ as the programming language.

@@

There is an optimal number of threads for each problem in each environment. This is the reason for the experiments ahead.

It is known that working concurrently opens up synchronization issues. But the {\it Multiprocessing} package (mp) offers both local and remote concurrency, effectively side-stepping the global interpreter lock by using sub-processes instead of threads. Because of this, the {\it Multiprocessing} module lets the programmer take full advantage of the fact that a machine has more than one processor, normally capable of 2 threads each.

The decision for {\it Multiprocessing} relies on the need to exchange information among processes (the pheromone trails or some information regarding load balancing, in our case), which would be harder to handle with thread-based parallelism. For this, we used the multiprocessing queue feature, that is thread safe.

Communication between processes can be achieved via {\it multiprocessing.Queue} and {\it multiprocessing.Pipe} which permits regular inter-process communication -IPC. (https://uwpce-pythoncert.github.io/Py300/threading-multiprocessing.html)

@@

For each red dot closest to the upper left corner, we call it {\it Utopia}, an impossible result with zero time duration and the best possible score.

\begin{figure}[H]
	\begin{center}
		\begin{tikzpicture}[scale=0.8, samples=100]
			\begin{axis}[
				myplotstyle,
				xlabel={\bf Scenario 1 computation times (s)},
				ylabel={\bf Score},
				]
				\addplot+[
				only marks,	
				nodes near coords = {\numProcs{}},
				visualization depends on = {value \thisrow{np} \as \numProcs},						
				]
				table[x=time, y=score, col sep=comma] {./csv/Shims_p1data20.csv};
			\end{axis}
			\fill [red] (canvas cs:x=0.1cm,y=5.6cm) circle (2pt);
		\end{tikzpicture}%
		~%
		%
		\begin{tikzpicture}[scale=0.8, samples=100]
			\begin{axis}[
				myplotstyle,
				xlabel={\bf Scenario 2 computation times (s)},
				]		
				\addplot+[
				only marks,	
				nodes near coords = {\numProcs{}},
				visualization depends on = {value \thisrow{np} \as \numProcs},						
				]
				table[x=time, y=score, col sep=comma] {./csv/Shims_p2data20.csv};
			\end{axis}
			\fill [red] (canvas cs:x=0.1cm,y=5.6cm) circle (2pt);		
		\end{tikzpicture}

@@

Observing Figure \ref{fig:mpShimsdata20}, it may be noticed that the best number of processes for {\it Scenario 1} (the closest to the {\it Utopia}) is \textbf{6}. For {\it Scenario 2}, the best number is \textbf{4}.

\begin{figure}[H]
	\begin{center}
		\begin{tikzpicture}[scale=0.8, samples=100]
			\begin{axis}[
				myplotstyle,
				xlabel={\bf Scenario 1 computation times (s)},
				ylabel={\bf Score},
				]
				\addplot+[
				only marks,	
				nodes near coords = {\numProcs{}},
				visualization depends on = {value \thisrow{np} \as \numProcs},						
				]
				table[x=time, y=score, col sep=comma] {./csv/Shims_p1data50.csv};
			\end{axis}
			\fill [red] (canvas cs:x=0.1cm,y=5.6cm) circle (2pt);
		\end{tikzpicture}%
		~%
		%
		\begin{tikzpicture}[scale=0.8, samples=100]
			\begin{axis}[
				myplotstyle,
				xlabel={\bf Scenario 2 computation times (s)},
				]
				\addplot+[
				only marks,	
				nodes near coords = {\numProcs{}},
				visualization depends on = {value \thisrow{np} \as \numProcs},						
				]
				table[x=time, y=score, col sep=comma] {./csv/Shims_p2data50.csv};
			\end{axis}
			\fill [red] (canvas cs:x=0.1cm,y=5.6cm) circle (2pt);
		\end{tikzpicture}
	\end{center}
	\caption{{\it mpShims} performance with $1.5 \times volume$.}
	\label{fig:mpShimsdata50}
\end{figure}

\begin{figure}[H]
	\begin{center}
		\begin{tikzpicture}[scale=0.8, samples=100]
			\begin{axis}[
				myplotstyle,
				xlabel={\bf Scenario 1 computation times (s)},
				ylabel={\bf Score},
				]
				\addplot+[
				only marks,	
				nodes near coords = {\numProcs{}},
				visualization depends on = {value \thisrow{np} \as \numProcs},						
				]
				table[x=time, y=score, col sep=comma] {./csv/ACO_p1data20.csv};
			\end{axis}
			\fill [red] (canvas cs:x=0.1cm,y=5.6cm) circle (2pt);
		\end{tikzpicture}%
		~%
		%
		\begin{tikzpicture}[scale=0.8, samples=100]
			\begin{axis}[
				myplotstyle,
				xlabel={\bf Scenario 2 computation times (s)},
				]
				\addplot+[
				only marks,	
				nodes near coords = {\numProcs{}},
				visualization depends on = {value \thisrow{np} \as \numProcs},						
				]
				table[x=time, y=score, col sep=comma] {./csv/ACO_p2data20.csv};
			\end{axis}
			\fill [red] (canvas cs:x=0.1cm,y=5.6cm) circle (2pt);
		\end{tikzpicture}
	\end{center}
	\caption{{\it mpACO} performance with $1.2 \times volume$.}
	\label{fig:mpACOdata20}
\end{figure}

@@

	\begin{center}
		\begin{tikzpicture}[scale=0.8, samples=100]
			\begin{axis}[
				myplotstyle,
				xlabel={\bf Scenario 1 computation times (s)},
				ylabel={\bf Score},
				]
				\addplot+[
				only marks,	
				nodes near coords = {\numProcs{}},
				visualization depends on = {value \thisrow{np} \as \numProcs},						
				]
				table[x=time, y=score, col sep=comma] {./csv/ACO_p1data50.csv};
			\end{axis}
			\fill [red] (canvas cs:x=0.1cm,y=5.6cm) circle (2pt);
		\end{tikzpicture}%
		~%
		%
		\begin{tikzpicture}[scale=0.8, samples=100]
			\begin{axis}[
				myplotstyle,
				xlabel={\bf Scenario 2 computation times (s)},
				]
				\addplot+[
				only marks,	
				nodes near coords = {\numProcs{}},
				visualization depends on = {value \thisrow{np} \as \numProcs},						
				]
				table[x=time, y=score, col sep=comma] {./csv/ACO_p2data50.csv};
			\end{axis}
			\fill [red] (canvas cs:x=0.1cm,y=5.6cm) circle (2pt);
		\end{tikzpicture}
	\end{center}
	\caption{{\it mpACO} performance with $1.5 \times volume$.}
	\label{fig:mpACOdata50}
\end{figure}

Observing figures \ref{fig:mpShimsdata20}, \ref{fig:mpShimsdata50}, \ref{fig:mpACOdata20}, \ref{fig:mpACOdata50}, it may be noticed that the best overall number of processes for all scenarios is \textbf{?}, because in average it is the number that most of the times is closer to the {\it Utopia} red bullet.

@@

		\multirow{2}{*}{\it {mpShims}} &\multirow{2}{*}{\bf {?}}& $f$              & 5.88        & 3.16    & 3.16         &          &          &         &   \\ 
		                               &                         &{\bf run-time (s)}& 1           & 10      & 41           &            &            &          &   \\ %299s
		\midrule[.1pt]

@@

        \multirow{2}{*}{\it {mpACO}}   &\multirow{2}{*}{\bf {?}}& $f$              & 5.82        & 3.12    &          &         &       &       &   \\ 
                                       &                         &{\bf run-time (s)}& 22          & 473     &           &         &       &       &   \\
